Action Groupoids and Fractional Types

### 2016-06-26 pibackground.tex
4.5
done with first pass over sec 4

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Operational Semantics and the Order of 1-Combinators}
\label{sec:pisem}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Semantics}
\label{sec:pisem}

@@

\begin{lemma}
\label{lem:distiterplus}
  For $p : \tau\iso\tau$, $m,n\in\Z$, we have a 2-combinator
  $\distiterplus{p}{m}{n} : (p^m \odot p^n) \isotwo p ^{m + n}$.
\end{lemma}

For an information-theoretic perspective on the language $\Pi$, we
think of a type containing $N$ values as an abstract system that
has~$N$ distinguishable states. According to the conventional theory
of information~\cite{Shannon1948}, the amount of information contained
in each state of a system with $N$ distinguishable states is $\log N$
bits of information. For example, the type $\mathsf{Bool}$ can be
thought of as an abstract system with two distinguishable states
labeled $\mathsf{true}$ and $\mathsf{false}$ each containing
$\log 2 = 1$ bit. Similarly, the type $\mathbb{3}$ can be thought of
as an abstract system with three distinguishable states each
containing $\log 3$ bits. The logarithmic map implies that information
contained in a composite state is the sum of the information contained
in its constituents. For example, the type
$\mathsf{Bool} \otimes \mathbb{3}$ can be thought of a composite
system consisting of two independent unrelated subsystems. Each state
of the composite system therefore contains
$\log (2 * 3) = \log 2 + \log 3$ bits which is the sum of the
information contained in each subsystem. Since all the combinators
preserve cardinality, they are also information-preserving.
